The Precambrian paleogeography of Laurentia

### 2020-11-07 Manuscript/Laurentia_Paleogeo_Manuscript.tex
update evaporite figure caption

Laurentia's growth continued into the Mesoproterozoic along the southeast margin through further juvenile terrane and arc accretion. Continental arc magmatism is interpreted to have occurred associated with the Pinwarian orogeny in the northeast Grenville province in Labrador from ca. 1.52 to 1.46 Ga \citep{Gower2002a}. Accretionary orogenesis recorded in the Grenville province includes accretion of the Quebecia composite arc terrane to Laurentia ca. 1.43 to 1.37 Ga \citep{Groulier2020a}. Far to the southwest along the margin, metamorphic rocks from Mesoproterozoic sedimentary and volcanic protoliths in northern New Mexico have been interpreted to indicate an interval of ca. 1.46 to 1.40 Ga orogenesis that has been named the Picuris orogeny \citep{Daniel2013a, Aronoff2016a}. In the midcontinent region,  deformation and metamorphism of post-Mazatzal orogeny sedimentary rocks is constrained to have occurred ca. 1.49 to 1.46 Ga associated with the Baraboo orogeny \citep{Medaris2003a, Holm2019a}. Picuris to Baraboo to Pinwarian orogenesis indicates convergent tectonism along the entirety of southeast Laurentia.  It was in this context of an active margin that there was major plutonism ca. 1.48 to 1.35 Ga that resulted in the A-type granitoids throughout both Mesoproterozoic and Paleoproterozoic provinces extending from the southwest United States up to the Central Gneiss Belt of Ontario to the northeast of Georgian Bay \citep{Slagstad2009a}. This plutonism has been interpreted to be associated with a combination of continental arc magmatism and melt generation within a back-arc region of Laurentia's long-live active margin \citep{Bickford2015a}. Younger magmatic activity (ca. 1.37 Ga) of the Southern Granite-Rhyolite Province suggests a similar active margin setting at that time \citep{Bickford2015a}. While an active margin interpretation for the Granite-Rhyolite Province, with arc and back-arc magmatism, has gained traction within the literature and is consistent with evidence for accretionary orogenesis in the Picuris, Baraboo and Pinware  orogens, the tectonic setting is often described as enigmatic given earlier interpretations of an anorogenic setting (see references in \citealp{Slagstad2009a}).

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{../Figures/Laurentia_evaporite_figure.pdf}
\caption{\textbf{Paleolatitude of Laurentia evaporites.} Left panel: The paleolatitude of evaporite deposits following the amalgamation of Laurentia as reconstructed by the simple Laurentia model shown in Fig. \ref{fig:Laurentia_paleolatitude} combined with the Phanerozoic model of \cite{Torsvik2017a} and as reconstructed by the model of \cite{Li2013a} for the Neoproterozoic. Proterozoic evaporite deposits in this panel are discussed in the text while Phanerozoic ones are taken from the compilation of \cite{Evans2006a}. The evaporite lines extend from the maximum to minimum age constraints while the points are at the preferred depositional age. The evaporite points are labeled with numbers both in the paleolatitude plot and the map of their present-day location in the right panel. The formations associated with these numbers are: 1 -- Altyn Formation (Belt Supergroup); 2: Wallace/Helena Formations -- Belt Supergroup; 3 -- Iqqittuq Formation (Bylot Supergroup); 4 -- Ten Stone Formation (Mackenzie Mountains Supergroup); 5 -- Minto Inlet Formation (Shaler Supergroup); 6 -- Redstone River Formation (Mackenzie Mountains Supergroup); 7 -- Kilian Formation (Shaler Supergroup); 8 -- Carboniferous Canadian Maritime; 9 -- Carboniferous Sverdrup; 10 -- Permian Midcontinental USA.}
\label{fig:Laurentia_evaporites}
\end{figure*}

>>>>

Laurentia's growth continued into the Mesoproterozoic along the southeast margin through further juvenile terrane and arc accretion. Continental arc magmatism is interpreted to have occurred associated with the Pinwarian orogeny in the northeast Grenville province in Labrador from ca. 1.52 to 1.46 Ga \citep{Gower2002a}. Accretionary orogenesis recorded in the Grenville province includes accretion of the Quebecia composite arc terrane to Laurentia ca. 1.43 to 1.37 Ga \citep{Groulier2020a}. Far to the southwest along the margin, metamorphic rocks from Mesoproterozoic sedimentary and volcanic protoliths in northern New Mexico have been interpreted to indicate an interval of ca. 1.46 to 1.40 Ga orogenesis that has been named the Picuris orogeny \citep{Daniel2013a, Aronoff2016a}. In the midcontinent region,  deformation and metamorphism of post-Mazatzal orogeny sedimentary rocks is constrained to have occurred ca. 1.49 to 1.46 Ga associated with the Baraboo orogeny \citep{Medaris2003a, Holm2019a}. Picuris to Baraboo to Pinwarian orogenesis indicates convergent tectonism along the entirety of southeast Laurentia.  It was in this context of an active margin that there was major plutonism ca. 1.48 to 1.35 Ga that resulted in the A-type granitoids throughout both Mesoproterozoic and Paleoproterozoic provinces extending from the southwest United States up to the Central Gneiss Belt of Ontario to the northeast of Georgian Bay \citep{Slagstad2009a}. This plutonism has been interpreted to be associated with a combination of continental arc magmatism and melt generation within a back-arc region of Laurentia's long-live active margin \citep{Bickford2015a}. Younger magmatic activity (ca. 1.37 Ga) of the Southern Granite-Rhyolite Province suggests a similar active margin setting at that time \citep{Bickford2015a}. While an active margin interpretation for the Granite-Rhyolite Province, with arc and back-arc magmatism, has gained traction within the literature and is consistent with evidence for accretionary orogenesis in the Picuris, Baraboo and Pinwarian orogens, the tectonic setting is often described as enigmatic given earlier interpretations of an anorogenic setting (see references in \citealp{Slagstad2009a}).

@@

\begin{figure*}
\centering
\includegraphics[width=7 in]{../Figures/Laurentia_evaporite_figure.pdf}
\caption{\textbf{Paleolatitude of Laurentia evaporites.} Left panel: The paleolatitude of evaporite deposits as reconstructed by the simple Laurentia model shown in Fig. \ref{fig:Laurentia_paleolatitude} combined with the Phanerozoic model of \cite{Torsvik2017a} and as reconstructed by the model of \cite{Li2013a} for the Neoproterozoic. Proterozoic evaporite deposits in this panel are discussed in the text while Phanerozoic ones are taken from the compilation of \cite{Evans2006a}. The evaporite lines extend from the maximum to minimum age constraints with points at the preferred depositional age. The evaporite paleolatitude points are labeled with numbers that correspond to the numbers on the present-day location map. These numbers correspond to: 1 -- Altyn Formation (Belt Supergroup); 2: Wallace/Helena Formations -- Belt Supergroup; 3 -- Iqqittuq Formation (Bylot Supergroup); 4 -- Ten Stone Formation (Mackenzie Mountains Supergroup); 5 -- Minto Inlet Formation (Shaler Supergroup); 6 -- Redstone River Formation (Mackenzie Mountains Supergroup); 7 -- Kilian Formation (Shaler Supergroup); 8 -- Silurian Michigan Basin ; 9 -- Devonian Western Canada; 10 -- Carboniferous Canadian Maritime; 11 -- Carboniferous Sverdrup; 12 -- Permian Midcontinental USA; 13 -- Jurassic Gulf of Mexico.}
\label{fig:Laurentia_evaporites}
\end{figure*}
